---
title: "Geospatial project "
output: html_document
date: "2025-03-05"
---
Downloading libraries
```{r}
rm(list = ls())

# Load libraries
library(sf)
library(tidyverse)
library(ggplot2)
library(leaflet)
library(osmdata)
library(raster)
library(sp)
library(ggspatial)
library(rasterVis)
```


Loading data about forest loss in Japan

```{r}

# Load Forest Loss (Raster)
forest_loss <- raster("/Users/admin/Desktop/Geospatial/Geospatial project/japan_forest_loss.tif")
```
Plotting forest loss




```{r}
plot(forest_loss)

```
```{r}
# Define color scheme
my_colors <- colorRampPalette(c("blue", "green", "yellow", "red"))

# High-quality raster plot
levelplot(forest_loss, col.regions = my_colors(100), main = "Forest Loss in Japan")
```


Loading data about urbanisation in Japan in 2000 and 2020 and plotting the urban change 
```{r}
# Load urban cover raster (2000 & 2023)
urban_2000 <- raster("/Users/admin/Desktop/Geospatial/Geospatial project/jpn_ppp_2000.tif")
urban_2020 <- raster("/Users/admin/Desktop/Geospatial/Geospatial project/jpn_ppp_2020.tif")

# Compute Urban Expansion
urban_change <- urban_2020 - urban_2000
```


```{r}
# Plot
levelplot(urban_2000, col.regions = my_colors(100), main = "Urban expansion from 2000 to 2020 in Japan")

```
```{r}
# Plot
levelplot(urban_2020, col.regions = my_colors(100), main = "Urban expansion from 2000 to 2020 in Japan")

```
```{r}
# Plot
levelplot(urban_change, col.regions = my_colors(100), main = "Urban expansion from 2000 to 2020 in Japan")

```


Loading the data about lakes and rivers in Japan
```{r}
library(sf)

# Load HydroLAKES for Japan (subset of global dataset)
hydrolakes <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/HydroLAKES_polys_v10_shp/HydroLAKES_polys_v10_shp/HydroLAKES_polys_v10.shp")

# Filter only lakes in Japan (assuming dataset has country codes)
japan_lakes <- hydrolakes[hydrolakes$Country == "Japan", ]
# Check if dataset is empty



```

We might not need hydroRIVERSbecause we will focul only on one main rivers later

```{r}

# Load HydroRIVERS for ASIA (subset of global dataset)
hydrorivres <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/HydroRIVERS_v10_as_shp/HydroRIVERS_v10_as_shp/HydroRIVERS_v10_as.shp")

# # Filter only lakes in Japan (assuming dataset has country codes)
# japan_rivers <- hydroriverss[hydrorivers$Country == "Japan", ]
# 

```

```{r}
# # Plot Asian Rivers
# Load Japan’s boundary polygon (Shapefile or GeoJSON)
japan_boundary <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/ne_10m_admin_0_countries/ne_10m_admin_0_countries.shp")
japan_boundary <- japan_boundary %>%
  filter(ISO_A3 == "JPN")  # Adjust column name if needed
```
These block of code also might not be needed as we don't use hydroRIVERS

```{r}
# Ensure both datasets have the same CRS
hydro_rivers <- st_transform(hydrorivres, crs = st_crs(japan_boundary))
# Perform spatial intersection (clip rivers by Japan boundary)
japan_rivers <- st_intersection(hydro_rivers, japan_boundary)

# Plot the extracted rivers
# Plot the extracted rivers
ggplot() +
  # Plot Japan Boundary
  geom_sf(data = japan_boundary, fill = NA, color = "black", size = 0.8) +
  
  # Plot Rivers
  geom_sf(data = japan_rivers, color = "blue", size = 0.6) +
  
  # Map Title & Theme
  labs(title = "Rivers in Japan") +
  theme_minimal()


```
```{r}
nat_earth_rivers <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/ne_50m_rivers_lake_centerlines/ne_50m_rivers_lake_centerlines.shp") 
japan_rivers_2 <- st_intersection(nat_earth_rivers, japan_boundary)
# Create the map with roads, railroads, and boundaries
```

Lakes and main river in Japan
```{r}
ggplot() +
  # Plot Japan Boundary
  geom_sf(data = japan_boundary, fill = NA, color = "black", size = 0.8) +
  
  # Plot Rivers
  geom_sf(data = japan_rivers_2, color = "blue", size = 0.6) +
  geom_sf(data = japan_lakes, color = "blue", size = 0.6) +
  # Map Title & Theme
  labs(title = "Rivers in Japan") +
  theme_minimal()

```


Loading data about protected areas 
```{r}
# Load the global WDPA dataset (Shapefile)
wdpa_global <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/WDPA_Mar2025_Public_shp/WDPA_Mar2025_Public_shp_0/WDPA_Mar2025_Public_shp-polygons.shp")

# Check the column names to find the country field
colnames(wdpa_global)

```
Filtering Japan 
```{r}
protected_areas_japan <- wdpa_global %>%
  filter(ISO3 == "JPN")  # Adjust column name if needed
```

Plotting the protected areas in japan 
```{r}
# Plot the filtered dataset
ggplot() +
  # Plot Japan Boundary
  geom_sf(data = japan_boundary, fill = NA, color = "black", size = 0.8) +
  
  # Plot Roads
  geom_sf(data = protected_areas_japan, color = "green", size = 0.6, alpha = 0.8) +
  
  # Map Title & Theme
  labs(title = "Protected areas in Japan") +
  theme_minimal()
```
Roads and Railroads in Japan 
```{r}
roads <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/ne_10m_roads/ne_10m_roads.shp")
japan_roads <- st_intersection(roads, japan_boundary)
railroads <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/ne_10m_railroads/ne_10m_railroads.shp")
japan_railroads <- st_intersection(railroads, japan_boundary)
```



```{r}
# Create the map with roads, railroads, and boundaries
ggplot() +
  # Plot Japan Boundary
  geom_sf(data = japan_boundary, fill = NA, color = "black", size = 0.8) +
  
  # Plot Roads
  geom_sf(data = japan_roads, color = "gray", size = 0.6, alpha = 0.8) +
  
  # Plot Railroads
  
  geom_sf(data = japan_railroads, color = "blue", size = 0.6, linetype = "dashed") +
  
  # Map Title & Theme
  labs(title = "Roads and Railroads in Japan") +
  theme_minimal()

```



```{r}
populated_places <- st_read("/Users/admin/Desktop/Geospatial/Geospatial project/ne_50m_populated_places/ne_50m_populated_places.shp")
japan_populated_places <- st_intersection(populated_places, japan_boundary)

```
```{r}
# Create the map with roads, railroads, and boundaries
ggplot() +
  # Plot Japan Boundary
  geom_sf(data = japan_boundary, fill = NA, color = "black", size = 0.8) +
  
  # Plot Roads
  geom_sf(data = japan_populated_places, color = "red", size = 0.4, alpha = 0.8) +
  
  # Map Title & Theme
  labs(title = "Populated places in Japan") +
  theme_minimal()
```
```{r}
# Find intersections
populated_protected_overlap <- st_intersection(japan_populated_places, protected_areas_japan)

# Plot affected areas
ggplot() +
  geom_sf(data = protected_areas_japan, fill = "green", alpha = 0.3) +
  geom_sf(data = populated_protected_overlap, fill = "red", alpha = 0.5) +
  labs(title = "Urban Expansion Over Protected Areas") +
  theme_minimal()

```
```{r}
protected_areas_japan <- st_make_valid(protected_areas_japan)
```


